---
title: "ADDS Final Report"
author: "Usmaan"
date: "2023-08-11"
output: html_document
---
```{r}
library(tidyverse)
library(ggplot2)
library(corrplot)
library(GGally)
```



```{r}
getwd()
setwd("C:/Users/Usmaan/Desktop/Final Project/Bank cutomer chirn DATASET")
getwd()
```
```{r}

d1=read.csv("C:/Users/Usmaan/Desktop/Final Project/Bank cutomer chirn DATASET/final.csv" , header = TRUE)
attach(d1)
```
```{r}
str(d1)
```
```{r}
d1$Geography<- as.factor(d1$Geography)
d1$Gender <- as.factor(d1$Gender)
d1$Satisfaction.Score <- as.factor(d1$Satisfaction.Score)
d1$HasCrCard<- as.factor(d1$HasCrCard)
d1$IsActiveMember <- as.factor(d1$IsActiveMember)
d1$Exited <- as.factor(d1$Exited)
d1$Complain <- as.factor(d1$Complain)
d1$NumOfProducts<-as.factor(d1$NumOfProducts)
d1$CreditScore<-as.numeric(d1$CreditScore)
d1$Age<-as.numeric(d1$Age)
d1$Tenure<-as.numeric(d1$Tenure)
d1$Balance<-as.numeric(d1$Balance)
d1$EstimatedSalary<-as.numeric(d1$EstimatedSalary)
d1$Point.Earned<-as.numeric(d1$Point.Earned)
```

```{r}
str(d1)
```
```{r}
num_vars <- names(d1)[sapply(d1, is.numeric)]
num_vars_summary <- summary(d1[num_vars])
num_vars_summary
```
```{r}
cat_vars <- names(d1)[sapply(d1, is.factor)]
cat_vars_summary <- summary(d1[cat_vars])
cat_vars_summary
```



```{r}
sum(is.na(d1))
```
```{r}
summary(d1)
```
```{r}
boxplot(Age)
```
```{r}
boxplot(Age)
boxplot(CreditScore)
boxplot(Balance)
boxplot(EstimatedSalary)
boxplot(Tenure)
boxplot(Point.Earned)
```

```{r}
gender.freq = table(d1$Gender)
pie(gender.freq, col = c("Red","Blue"), main = "Pie Chart For Gender")
```
# To visualize the gender in percentage
```{r}
gender_percentages <- table(d1$Gender) / nrow(d1) * 100

gender_data <- data.frame(
  gender = names(gender_percentages),
  percentage = gender_percentages
)
ggplot(data = gender_data, aes(x = "", y = gender_percentages, fill = gender)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Gender Percentage Pie Chart") +
  theme_void() +   geom_text(aes(label = paste0(round(gender_percentages, 1), "%")), position = position_stack(vjust = 0.5))

```
# Count of gender
```{r}
table(d1$Gender)
```
# To Count and visualize the churn in pie chart
```{r}
table(d1$Exited)
```
```{r}
exited.f <- factor(Exited , levels = c(0,1), labels = c('Not Churn','Churn'))


exited_count <- table(exited.f) / nrow(d1)  * 100

churn_data <- data.frame(
  types = names(exited_count),
  percentage = exited_count
)
ggplot(data = churn_data, aes(x = "", y = exited_count, fill = types)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Churn Percentage in Pie Chart") +
  theme_void() +   geom_text(aes(label = paste0(round(exited_count, 1), "%")), position = position_stack(vjust = 0.5))
```
```{r}
ggplot(data=d1)+geom_bar(mapping=aes(x= exited.f , fill = exited.f)) + labs(title = "Count of Churn" , y = "Count" , x = "Churn or Not") + theme_dark()
```
# Geoagraphy Percentage
```{r}
geography_percentages <- table(d1$Geography) / nrow(d1) * 100

geography_data <- data.frame(
  geography = names(geography_percentages),
  percentage = geography_percentages
)
ggplot(data = geography_data, aes(x = "", y = geography_percentages, fill = geography)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "Geography Percentage Pie Chart(EUROPE)") +
  theme_void() +   geom_text(aes(label = paste0(round(geography_percentages, 1), "%")), position = position_stack(vjust = 0.5))
```
```{r}
table(d1$Geography)
```
# To show number of products
```{r}
table1=table(d1$NumOfProducts)
table1<- as.data.frame(table1)
colnames(table1)<- c("NumOfProducts", "Frequency")
table1
#row.names(table1) <- c("One Product", "Two Products", "Three Products", "Four Products")
#attr(table1, "title") <- "Frequency of Number of Products"
```
```{r}
ggplot(data=d1)+geom_bar(mapping=aes(x= NumOfProducts , fill= NumOfProducts ))
```
# Count of Credit card customer.
```{r}
cr.f <- factor(HasCrCard , levels = c(0,1), labels = c("Don't Have Credit Card","Have Credit Card"))

ggplot(data=d1)+geom_bar(mapping=aes(x= cr.f , fill = exited.f),position = "dodge") + labs(title = "Count of Credit Card Customers" , y = "Count" , x = "HasCrCard") + theme_dark()
```
```{r}
cr_table=table(d1$HasCrCard,d1$Exited)
tcr<- as.data.frame(cr_table)
colnames(tcr)<- c("HasCrCard","Churn","Count")
tcr
```






```{r}
table2=table(cr.f)
table2<- as.data.frame(table2)
colnames(table2)<- c("HasCrCard", "Frequency")
table2
```

# Count of active member
```{r}
active.f <- factor(IsActiveMember , levels = c(0,1), labels = c("Not Active","Active"))
ggplot(data=d1)+geom_bar(mapping=aes(x= active.f , fill = exited.f), position='dodge') + labs(title = "Count of Active Customers" , y = "Count" , x = "IsActiveMember") + theme_dark()
```
```{r}
active_table=table(d1$IsActiveMember,d1$Exited)
tac<- as.data.frame(active_table)
colnames(tac)<- c("IsActive","Churn","Count")
tac
```



```{r}
table3=table(active.f)
table3<- as.data.frame(table3)
colnames(table3)<- c("IsActiveMember", "Frequency")
table3
```

# Count of customers who complained to bank and not complained
```{r}
complain.f <- factor(Complain , levels = c(0,1), labels = c("Not Complained","Complained"))
ggplot(data=d1)+geom_bar(mapping=aes(x= complain.f , fill = exited.f),position = 'dodge') + labs(title = "Count of Customers Who Complained to Bank" , y = "Count" , x = "Complain") + theme_dark()
```
```{r}
gender_occupation_table=table(d1$Complain,d1$Exited)
t2<- as.data.frame(gender_occupation_table)
colnames(t2)<- c("Complain","Churn","Count")
t2

```




```{r}
table4=table(complain.f)
table4<- as.data.frame(table4)
colnames(table4)<- c("Complain", "Frequency")
table4
```

# Satisfiction Score Count
```{r}
table5=table(Satisfaction.Score)
table5<- as.data.frame(table5)
colnames(table5)<- c("Satisfaction Level", "Frequency")
table5
```
```{r}
ggplot(data=d1)+geom_bar(mapping=aes(x= Satisfaction.Score , fill = exited.f), position = 'dodge') + labs(title = "Satisfaction Score Count" , y = "Count" , x = "Satisfaction.Score") + theme_classic()


satis_table=table(Satisfaction.Score,Exited)
tss<- as.data.frame(satis_table)
colnames(tss)<- c("Satisfaction Score","Churn","Count")
tss
```
### Numerical Data Visualization 

```{r}
boxplot(Tenure~exited.f,data = d1)

ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y=Tenure,fill= Gender , alpha = 0.75)) + labs(title='Tenure vs Churn',x = 'Churn or Not' , y = 'Tenure') + theme_classic()

ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y=Tenure,fill= exited.f , alpha = 0.75)) + labs(title='Tenure vs Churn',x = 'Churn or Not' , y = 'Tenure') + theme_classic()

table(d1$Gender,d1$Tenure,d1$Exited)
```
```{r}
ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y=Balance,fill= exited.f , alpha = 0.75)) + labs(title='Balance vs Churn',x = 'Churn or Not' , y = 'Balance') + theme_classic()

summary(d1$Balance)
```
```{r}

ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y=EstimatedSalary,fill= exited.f , alpha = 0.75)) + labs(title='Salary vs Churn',x = 'Churn or Not' , y = 'Salary') + theme_classic()
```

```{r}
ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y=Point.Earned,fill= exited.f , alpha = 0.75)) + labs(title='Earned points vs Churn',x = 'Churn or Not' , y = 'Earned Points') + theme_classic()

ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y=Point.Earned,fill= exited.f , alpha = 0.75)) + labs(title='Earned points vs Churn',x = 'Churn or Not' , y = 'Earned Points') + theme_classic() +  geom_text(aes(x = exited.f, y = Point.Earned, label = Point.Earned),position =  position_dodge(width = 0.75), vjust = 5 , hjust = 2)
```

```{r}
summary(d1$Point.Earned)
```
```{r}
boxplot(Age)
ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y= Age,fill= exited.f , alpha = 0.75)) + labs(title='Age vs Churn',x = 'Churn or Not' , y = 'Age') + theme_classic()
```

```{r}
summary(d1$Age)
```



```{r}
boxplot(CreditScore)
ggplot(data=d1)+geom_boxplot(mapping=aes(x=exited.f,y= CreditScore,fill= exited.f , alpha = 0.75)) + labs(title='Credit Score vs Churn',x = 'Churn or Not' , y = 'Scored Credit Points') + theme_classic()
```











### Numerical Variables

# Distribution of credit score
# bw = higher value will decrease the resolution = density plot
# adjust = 
```{r}
breaks1 <- seq(0, 100, by = 20)
data_ranges1 <- cut(d1$Age, breaks = breaks1, include.lowest = TRUE,labels = c("0-20","20-40","40-60","60-80","80-100"))
table6=table(data_ranges1)
table6<- as.data.frame(table6)
colnames(table6)<- c("Age Range", "Frequency")
table6
#if you want to determine in which range 200 will fall based on these breakpoints, you can think of the ranges as intervals. Each range interval is defined by two consecutive breakpoints. For example, the interval between 100 and 200 would be represented as (100, 200], where the square bracket ( indicates that the range doesn't include the left endpoint (100), and the square bracket ] indicates that the range includes the right endpoint (200).So, in the context of the generated sequence, the range (100, 200] includes values greater than 100 and less than or equal to 200. Therefore, 200 falls within this range.
```


```{r}
ggplot(data=d1)+geom_histogram(mapping=aes(x=CreditScore),binwidth = 25,color="blue",fill="lightgreen")

ggplot(data=d1)+geom_histogram(mapping=aes(x=CreditScore,y=..density..),binwidth = 25,color="blue",fill="lightgreen")+geom_density(mapping=aes(x=CreditScore), color= 'red')+theme_classic()

ggplot(data=d1)+geom_histogram(mapping=aes(x=CreditScore,y=..density..),binwidth = 100,color="blue",fill="lightgreen")+geom_density(mapping=aes(x=CreditScore), color= 'red')+theme_classic()
```
```{r}
breaks2 <- seq(300, 900, by = 100)
data_ranges2 <- cut(d1$CreditScore, breaks = breaks2, include.lowest = TRUE,labels = c("300-400","400-500","500-600","600-700","700-800","800-900"))
table7=table(data_ranges2)
table7<- as.data.frame(table7)
colnames(table7)<- c("Score Range", "Frequency")
table7
```

```{r}
ggplot(data=d1)+geom_histogram(mapping=aes(x=Point.Earned),binwidth = 50,color="blue",fill="lightgreen")

ggplot(data=d1)+geom_histogram(mapping=aes(x=Point.Earned,y=..density..),binwidth = 50,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Point.Earned),color="red")

ggplot(data=d1)+geom_histogram(mapping=aes(x=Point.Earned,y=..density..),binwidth = 100,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Point.Earned),color="red")
```
```{r}
breaks3 <- seq(100, 1000, by = 100)
data_ranges3 <- cut(d1$Point.Earned, breaks = breaks3, include.lowest = TRUE,labels = c("100-200","200-300","300-400","400-500","500-600","600-700","700-800","800-900","900-1000"))
table8=table(data_ranges3)
table8<- as.data.frame(table8)
colnames(table8)<- c("Point Range", "Frequency")
table8
```



```{r}
ggplot(data=d1)+geom_histogram(mapping=aes(x=Age),binwidth = 2,color="blue",fill="lightgreen")

ggplot(data=d1)+geom_histogram(mapping=aes(x=Age,y=..density..),binwidth = 2,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Age),color="red")

ggplot(data=d1)+geom_histogram(mapping=aes(x=Age,y=..density..),binwidth = 10,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Age),color="red")

```
```{r}
breaks1 <- seq(0, 60, by = 10)
data_ranges1 <- cut(d1$Age, breaks = breaks1, include.lowest = TRUE,labels = c("0-10","10-20","20-30","30-40","40-50","50-60"))
table6=table(data_ranges1)
table6<- as.data.frame(table6)
colnames(table6)<- c("Age Range", "Frequency")
table6
```

```{r}
ggplot(data=d1)+geom_histogram(mapping=aes(x=Tenure),binwidth = 1,color="blue",fill="lightgreen")

ggplot(data=d1)+geom_histogram(mapping=aes(x=Tenure,y=..density..),binwidth = 1,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Tenure),color="red")

ggplot(data=d1)+geom_histogram(mapping=aes(x=Tenure,y=..density..),binwidth = 2,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Tenure),color="red")
```
```{r}
breaks4 <- seq(0, 10, by = 2)
data_ranges4 <- cut(d1$Tenure, breaks = breaks4, include.lowest = TRUE,labels = c("0-2","2-4","4-6","6-8","8-10"))
table9=table(data_ranges4)
table9<- as.data.frame(table9)
colnames(table9)<- c("Tenure Range", "Frequency")
table9
```

```{r}
scientific_to_decimal <- function(x) {
  format(x, scientific = FALSE)
} # to convert scientific value to decimal, convert scientific notation to decimal notation on the X-axis of a histogram

ggplot(data=d1)+geom_histogram(mapping=aes(x=Balance),binwidth = 10000,color="blue",fill="lightgreen")+ scale_x_continuous(labels = scientific_to_decimal)

ggplot(data=d1)+geom_histogram(mapping=aes(x=Balance,y=..density..),binwidth = 10000,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Balance),color="red")+ scale_x_continuous(labels = scientific_to_decimal)

ggplot(data=d1)+geom_histogram(mapping=aes(x=Balance,y=..density..),binwidth = 50000,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=Balance),color="red")+ scale_x_continuous(labels = scientific_to_decimal)


```
```{r}
breaks5 <- seq(0,300000, by =50000)
data_ranges5 <- cut(d1$Balance, breaks = breaks5, include.lowest = TRUE,labels = c("0-50K","50K-100K","100K-150K","150K-200K","200K-250K","250K-300K"))
table10=table(data_ranges5)
table10<- as.data.frame(table10)
colnames(table10)<- c("Euro Range", "Frequency")
table10
```


```{r}
ggplot(data=d1)+geom_histogram(mapping=aes(x=EstimatedSalary),binwidth = 10000,color="blue",fill="lightgreen")

ggplot(data=d1)+geom_histogram(mapping=aes(x=EstimatedSalary,y=..density..),binwidth = 10000,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=EstimatedSalary),color="red")

ggplot(data=d1)+geom_histogram(mapping=aes(x=EstimatedSalary,y=..density..),binwidth = 50000,color="blue",fill="lightgreen")+ theme_classic()+geom_density(mapping=aes(x=EstimatedSalary),color="red")
```
```{r}
breaks6 <- seq(0,200000, by =50000)
data_ranges6 <- cut(d1$EstimatedSalary, breaks = breaks6, include.lowest = TRUE,labels = c("0-50K","50K-100K","100K-150K","150K-200K"))
table11=table(data_ranges6)
table11<- as.data.frame(table11)
colnames(table11)<- c("Euro Range", "Frequency")
table11
```
### Correlation Plot
```{r}
str(d1)
```
```{r}
d2corplot <- select(d1, CreditScore, Age, Tenure, Balance, EstimatedSalary, Point.Earned)
d2corplot

str(d2corplot)
```


```{r}
cor(d2corplot)

```
```{r}
GGally::ggcorr(d2corplot) + theme_dark()
```
```{r}
fx= select(d1, -(CustomerId:Surname))
fx
```
```{r}
str(fx)
```
```{r}
z1=write.csv(fx,file="EditedPYT.csv",row.names = F)
getwd()
```

```{r}
tst1= filter(d1,CustomerId<15691000)
tst1
```
```{r}
zz1=write.csv(tst1,file="5009.csv",row.names = F)
getwd()
```

###Correlation Analysis

```{r}
o2=lm((Point.Earned)~Balance,data=d1)
summary(o2)
```
```{r}
cor.test(d1$Point.Earned,d1$Balance,conf.level = 0.95)
```
```{r}
cor(d1$Point.Earned,d1$Balance)
```
```{r}
ggplot(data = d1) + geom_point(mapping = aes(x= Balance, y =Point.Earned),color="blue")
```
```{r}
o3=aov(Balance ~ IsActiveMember, data = d1)
o3
```


```{r}
summary(o3)
```
















